Tricolorings of Knots

### 2018-04-24 presentation/presentation.tex
Added periods

\begin{frame}
  \frametitle{Knot Definition}
  \begin{itemize}
  \item A \textit{Knot} is a simple closed curve in $S^3$ ($\mathbb{R}^3\cup \{\infty\}$)
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Equivalence}
  \begin{itemize}
  \item All knots are homeomorphic to $S^1$.
    \pause
  \item Two knots $K$ and $J$ are \textit{Ambient Isotopic}
    if there is a homotopy $H:S^3\times[0,1]\rightarrow S^3$ such that
    \begin{itemize}
      \pause
    \item $H(x,0)$ is the identity on $S^3$
      \pause
    \item The image $H(K,1)$ is $J$
      \pause
    \item For each fixed $t\in[0,1]$ the map $H(x,t)$ is a
      homeomorphism.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Diagrams}
  \begin{itemize}
  \item A \textit{diagram} $D$ of a knot $K$ is a projection of
    $K$ to a plane such that
    \begin{itemize}
      \pause
    \item At most two points can be projected to the same point on the plane.
      \pause
    \item All crossings must be either overcrossings or undercrossings.
      \pause
    \item We preserve over and under crossing information      
    \end{itemize}
    \begin{center}
      \visible<3->{\includegraphics[scale=.2]{bad-crossing}}
    \end{center}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Tricoloring}
  \begin{itemize}
  \item A \textit{tricoloring} of a diagram is an assignment
    of one of three colors to each arc such that for all crossings
    \begin{itemize}
      \pause
    \item Only a single color is present
      \pause
    \item All three colors are present
    \end{itemize}
    \pause
  \item A tricoloring is \textit{nontrivial} if at least two colors are used
    \pause
  \item \textbf{Theorem:} The existence of a non-trivial tricoloring is a knot
    invariant. Such a knot is called tricolorable (or $3$-colorable).
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Knot Group}
  \begin{itemize}
  \item Since all knots are homeomorphic to $S^1$ their
    fundamental groups are isomorphic to $\mathbb{Z}$.
    \pause
  \item Given a knot $K$, the \textit{knot group} is
    the fundamental group of the knots complement  $\pi_1(S^3\setminus K,*)$
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Wirtinger Presentation}
  \begin{itemize}
  \item For other knots we use a tool called the \textit{Wirtinger Presentation}.
    \pause
  \item Let $K$ be a knot and $D$ a diagram for $K$. Choose an orientation of $D$.
    Let $x_1,\ldots, x_n$ denote the arcs of $D$ and, for each crossing, give a
    relation $r_1,\ldots, r_m$ such that $r_i\equiv (x_i x_j x_i^{-1}=x_k)$ where, when approaching
    $x_k$ is the arc leaving the crossing, the top arc is $x_i$ and the bottom arc entering
    is $x_j$.
    \pause
  \item Then the knot group of $K$ has presentation
    \[
      \pi_1(S^3\setminus K) \equiv \langle x_1,\ldots, x_n| r_1,\ldots r_m\rangle
    \]
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Knot Group of the Trefoil}
  \begin{center}
    \includegraphics[scale=.25]{annotated-trefoil}
  \end{center}
  \begin{itemize}
  \item From this diagram the knot group of the trefoil is 
    \[
      \langle x_1,x_2,x_3|x_1x_2x_1^{-1}=x_3,x_3x_1x_3^{-1}=x_2,x_2x_3x_2^{-1}=x_1\rangle
    \]\pause
  \item Which simplifies to $\langle x_1,x_2|x_1 x_2 x_1= x_2 x_1 x_2\rangle$
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Knot Definition}
  \begin{itemize}
  \item A \textit{Knot} is a simple closed curve in $S^3$ ($\mathbb{R}^3\cup \{\infty\}$).
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Equivalence}
  \begin{itemize}
  \item All knots are homeomorphic to $S^1$.
    \pause
  \item Two knots $K$ and $J$ are \textit{Ambient Isotopic}
    if there is a homotopy $H:S^3\times[0,1]\rightarrow S^3$ such that:
    \begin{itemize}
      \pause
    \item $H(x,0)$ is the identity on $S^3$.
      \pause
    \item The image $H(K,1)$ is $J$.
      \pause
    \item For each fixed $t\in[0,1]$ the map $H(x,t)$ is a
      homeomorphism.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Diagrams}
  \begin{itemize}
  \item A \textit{diagram} $D$ of a knot $K$ is a projection of
    $K$ to a plane such that:
    \begin{itemize}
      \pause
    \item At most two points can be projected to the same point on the plane.
      \pause
    \item All crossings must be either overcrossings or undercrossings.
      \pause
    \item We preserve over and under crossing information.
    \end{itemize}
    \begin{center}
      \visible<3->{\includegraphics[scale=.2]{bad-crossing}}
    \end{center}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Tricoloring}
  \begin{itemize}
  \item A \textit{tricoloring} of a diagram is an assignment
    of one of three colors to each arc such that for all crossings:
    \begin{itemize}
      \pause
    \item Only a single color is present.
      \pause
    \item All three colors are present.
    \end{itemize}
    \pause
  \item A tricoloring is \textit{nontrivial} if at least two colors are used.
    \pause
  \item \textbf{Theorem:} The existence of a non-trivial tricoloring is a knot
    invariant. Such a knot is called tricolorable (or $3$-colorable).
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Knot Group}
  \begin{itemize}
  \item Since all knots are homeomorphic to $S^1$ their
    fundamental groups are isomorphic to $\mathbb{Z}$.
    \pause
  \item Given a knot $K$, the \textit{knot group} is
    the fundamental group of the knots complement  $\pi_1(S^3\setminus K,*)$.
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Wirtinger Presentation}
  \begin{itemize}
  \item For other knots we use a tool called the \textit{Wirtinger Presentation}.
    \pause
  \item Let $K$ be a knot and $D$ a diagram for $K$. Choose an orientation of $D$.
    Let $x_1,\ldots, x_n$ denote the arcs of $D$ and, for each crossing, give a
    relation $r_1,\ldots, r_m$ such that $r_i\equiv (x_i x_j x_i^{-1}=x_k)$ where, when approaching
    $x_k$ is the arc leaving the crossing, the top arc is $x_i$ and the bottom arc entering
    is $x_j$.
    \pause
  \item Then the knot group of $K$ has presentation
    \[
      \pi_1(S^3\setminus K) \equiv \langle x_1,\ldots, x_n| r_1,\ldots r_m\rangle.
    \]
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Knot Group of the Trefoil}
  \begin{center}
    \includegraphics[scale=.25]{annotated-trefoil}
  \end{center}
  \begin{itemize}
  \item From this diagram the knot group of the trefoil is 
    \[
      \langle x_1,x_2,x_3|x_1x_2x_1^{-1}=x_3,x_3x_1x_3^{-1}=x_2,x_2x_3x_2^{-1}=x_1\rangle.
    \]\pause
  \item Which simplifies to $\langle x_1,x_2|x_1 x_2 x_1= x_2 x_1 x_2\rangle$.
  \end{itemize}
\end{frame}
